---
title: "Reading, inspecting, and writing data"
author: "Jerid Francom"
date-modified: today
format: html
bibliography: [bibliography.bib, packages.bib]
biblio-style: apalike
link-citations: true
---

# Overview

## About the data

```{r}
#| label: load-packages

# Load packages
library(readr)   # for reading and writing data
library(dplyr)   # for data manipulation
```

1. What is the name of the data source?

The data for this lab comes from the [Manually-Annotated Subcorpus (MASC)](https://anc.org/data/masc/about/) [@Ide2013]. 

The data origin is seen below: 

```{r}
#| label: read-data-origin
#| echo: false

# Read and print the data origin file
read_csv("data/original/data_origin.csv")
```

2. Where did it come from?

The MASC is a annotated subset of the American National Corpus.

3. What is the sampling frame?

The corpus is of American English covering a wide range of genres since the 1990s.

## Inspect the data

The data dictionary is seen below:

```{r}
#| label: read-data-dictionary
#| echo: false
#| message: false

# Read and print the data dictionary
read_csv("data/derived/data_dictionary.csv")
```

Here is a table version of the data dictionary in Table \@ref(tab:data-dictionary).

```{r}
#| label: data-dictionary
#| echo: false
#| message: false

# Load packages
library(knitr)   # for tables

# Read and print the data dictionary
read_csv("data/derived/data_dictionary.csv") |>
  kable(caption = "Data dictionary for the MASC data.")
```

1. How many variables are there? 

From the data dictionary, we can see that there are 9 variables.

2. What are the variable types?

From the data dictionary, we can see that there are 2 numeric variables and 7 character variables.

```{r}
#| label: read-data
#| echo: false
#| message: false

# Read the data
masc_df <- read_csv("data/derived/data.csv")
```

After loading the dataset into R, we can see that the dataset has almost 600k observations (591,036). 

```{r}
#| label: inspect-data

# Preview
glimpse(masc_df)
```

## Subset the data

In this section, I will subset the data to only include the columns `modality`, `genre`, `term`, and `pos` for only adjectives (`JJ`, `JJR`, and `JJS`).

```{r}
#| label: subset-data

# Subset the data
masc_adj_df <- 
  masc_df |> # pipe the data into the next function
  select(modality, genre, term, pos) |> # select the columns
  filter(pos %in% c("JJ", "JJR", "JJS")) # filter the rows

# Preview
glimpse(masc_adj_df)
```

1. How does the subsetted data differ from the original data?

The subsetted data only contains adjectives, while the original data contained all parts of speech. And now there is only a subset of the columns.

2. What are the dimensions of the subsetted data?

The subsetted data has 4 columns and 33,994 observations. The original data had 9 columns and 591,036 observations. 

3. What are the variable types?

All of the variables are character variables.

## Write the data

I will write the `masc_adj_df` data to a csv file with the name `masc_adj.csv`. It will add it to the `data/derived` folder.

```{r}
#| label: write-data

# Write the data
write_csv(masc_adj_df, "data/derived/masc_adj.csv")
```

## Self-assessment

In this lab I learned how to read, inspect, subset, and write data. I also learned how to use the pipe operator (`|>`) to make my code more readable. 

I found reading and creating file paths to be the most challenging part of this lab. I also found the pipe operator to be a little confusing at first, but I think I am starting to get the hang of it.